---
title: "Introduction to Concstats"
author: "Andreas Schneider"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: TRUE
    toc_depth: 3
vignette: >
  %\VignetteIndexEntry{Introduction to Concstats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      fig.align = "center",
                      warning = FALSE,
                      screenshot.force = FALSE)
```


```{R,echo=FALSE,results='hide',message=FALSE,include=FALSE, eval=TRUE}
if (!require(knitr))
  install.packages("knitr", repos = "http://cran.us.r-project.org")
if (!require(devtools))
  install.packages("devtools", repos = "http://cran.us.r-project.org")
if (!require(dplyr))
  install.packages("dplyr", repos = "http://cran.us.r-project.org")
if (!require(kableExtra))
  install.packages("kableExtra", repos = "http://cran.us.r-project.org")
if (!require(ggplot2))
  install.packages("ggplot2", repos = "http://cran.us.r-project.org")
```

The goal of the ``concstats`` package is to offer a set of alternative and/or
additional measures to better determine a given market structure and therefore
reduce uncertainty with respect to a given market situation. Various functions
or groups of functions are available to achieve the desired goal. 

## Installation
To install the latest development version of `concstats` directly from [GitHub](https://github.com/schneiderpy/concstats) use: 

```{r, message=FALSE, warning=FALSE, results = "hide", eval=FALSE}
library(devtools) #  Tools to Make Developing R Packages Easier 
devtools::install_github("schneiderpy/concstats")
```

Then, load the package.
```{r, message=FALSE, warning=FALSE}
library(concstats)
```

## Data
The following examples use mainly fictitious data to present the functions.
However, if you want to test the functionality in more detail, the package comes
with a small data set of real paraguayan credit cooperatives (creditcoops). There
are 22 paired observations for real paraguayan credit cooperatives 
(with assets > 11 Mio. USD) for 2016 and 2018 with their respective total loans 
granted. For a better visualization there is an additional column with the 
transformed total loans. For further information on the data please see the
`data` help file.

```{r data, echo=FALSE}
data("creditcoops")
head(creditcoops)
```

## Group wrapper overview
At the moment, there are the following groups of functions available:    
- `mstruct` is a wrapper for market structure measures    
- `inequ` is a wrapper for inequality and diversity measures    
- `comp` is a wrapper with different concentration measures    
- `concstats` is a function which calculates a set of preselected measures in a 
    one step procedure to get a quick overview of a given market structure    

The functions will be presented in more details in the following short
step-by-step guide. 

## Examples
We will use a vector which represents market participants with their respective 
market shares (in decimal form): 

### mstruct
The wrapper includes the following individual functions: ```firm```, ```nrs_eq```,
```top```, ```top3```, ```top5```, and ```all```
```{r examples-top5, echo=TRUE}
test_share <- c(0.35, 0.4, 0.05, 0.1, 0.06, 0.04, 0, 0)
test_share_top5 <- top5(test_share) # top 5 market share
```
You should have noticed that the market shares are in decimal form. There are 
eight market participants, however, two have no market shares, by default 
```concstats``` treats 0 as NA. The result is a top 5 market share of 96 %.
You can also access each function through their respective argument in the 
group wrapper:

```{r examples-top, echo=TRUE}
test_share <- c(0.35, 0.4, 0.05, 0.1, 0.06, 0.04, 0, 0)
test_share_top <- mstruct(test_share, type = "top") # top market share
```

Or, just calculate all measures of the group wrapper.
```{r examples-mstruct-all, echo=TRUE}
test_share <- c(0.35, 0.4, 0.05, 0.1, 0.06, 0.04, 0, 0)
test_share_mstruct <- mstruct(test_share, type = "all")
```

The result is a table with the market structure measures.

### inequ
The inequality and diversity group includes the functions ```entropy```, 
```gini```, ```berry```, ```palma```, ```grs```, and ```all```.

```{r examples-entropy, echo=TRUE}
test_share <- c(0.35, 0.4, 0.05, 0.1, 0.06, 0.04, 0, 0)
test_share_entropy <- entropy(test_share)

# and as a standardized value
test_share_entropy2 <- entropy(test_share, unbiased = TRUE)
```

### comp
The group wrapper for competition measures includes the functions ```hhi```, 
```hhi_d```, ```hhi_min```, ```dom```, ```sten```, and ```all```

```{r examples-hhi, echo=TRUE}
test_share <- c(0.35, 0.4, 0.05, 0.1, 0.06, 0.04, 0, 0)
test_share_hhi <- hhi(test_share)

# a standardized value
test_share_hhi2 <- hhi(test_share, unbiased = TRUE)

# the min average of the hhi
test_share_hhi3 <- comp(test_share, unbiased = TRUE, type = "hhi_min")
```

### concstats
A single function which calculates a set of eight preselected measures in a 
one step procedure for a first overview of a given market structure.

```{r examples-concstats, echo=TRUE}
test_share <- c(0.35, 0.4, 0.05, 0.1, 0.06, 0.04, 0, 0)
test_share_conc <- concstats(test_share)
```

## Visualization
The scope of the package is to calculate market structure and concentration 
measures to get a quick and more informed overview of a given market situation.
However, it is good practice to visualize your data in an exploratory step or in
reporting your results. The package ```concstats``` works fine with other EDA or 
data visualization packages e.g. [```overviewR```](https://CRAN.R-project.org/package=overviewR),
[```dataexplorer```](https://CRAN.R-project.org/package=DataExplorer), [```kableExtra```](https://CRAN.R-project.org/package=kableExtra) or [```ggplot2```](https://CRAN.R-project.org/package=ggplot2) to name a view.

Some examples how you can accomplish this.
Let us assume one would like to use the group measure for e.g. market structure,
and keep the resulting table. We can refine the table using ```kableExtra``` 
which works nice with ```knitr```.

This time, we will use our ```creditcoops``` data set, which comes with the 
package.

```{r visualization-data, message=FALSE, echo=TRUE}
data("creditcoops")
head(creditcoops)
```

You will need the following two packages.
```{r visualization-prep}
library(dplyr)
library(kableExtra)
```

Now, we will filter out data for the year 2016.
```{r visualization-tab, echo=TRUE}
coops_2016 <- creditcoops %>% dplyr::filter(year == 2016)
head(coops_2016)

coops_2016 <-  coops_2016[["total_loans"]]
# atomic vector of total loans

# We then use the new object `coops_2016` to calculate the market structure measures
# as a group in a one-step-procedure:
coops_2016_mstruct <- mstruct(coops_2016, type = "all")
coops_2016_mstruct_tab <- coops_2016_mstruct %>% 
  kableExtra::kbl(caption = "Market structure 2016", digits = 2,
                  booktabs = T, align = "r") %>% 
  kableExtra::kable_classic(full_width = F, html_font = "Arial")
coops_2016_mstruct_tab
```

The result is a nice reusable table.

Now, let's go a step further. We will compare the two samples for 2016 and 2018.
For this purpose, we will select from our `creditcoops` data set the relevant
columns (coop_id, year, paired, and total_loans_log) and make a new data frame.
```{r visualization-sample-prep, echo=FALSE}
df_shares <- creditcoops %>% dplyr::select(coop_id, year, paired, total_loans_log)
```
Make sure you have installed the `ggplot2` package. Load the package.
```{r visualization-ggplot}
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
```

```{r visualization-sample-plot, echo=FALSE, fig.height=5, fig.width=5}
df_shares_plot <- df_shares %>%
  ggplot2::ggplot(aes(year, total_loans_log, fill = year)) +
  ggplot2::geom_boxplot() +
  ggplot2::geom_point() +
  ggplot2::geom_line(aes(group = paired)) +
  ggplot2::labs(title = "Credit cooperatives (type A)", y = "Total loans (log)",
       caption = "Source: Andreas Schneider with data from INCOOP") +
  ggplot2::theme(legend.position = "none")
df_shares_plot
```

Having a look a the output, we see a box plot with paired values (cooperatives)
for the two sample years 2016 and 2018.


