---
title: "concstats"
description: >
  Learn how to easily use the functionallity of `concstats`.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{concstats}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(concstats)
```
The easiest situation is, that you have already a vector with the calculated market share of each market participant.

### Example:

```{r, eval=FALSE}
df$share # a vector of a data frame with the calculated market share (in decimal form)
share <- c(0.4, 0.226, 0.334) # a single vector
```

However, most of the time you have a data frame with just the values (in local currency) of each market participant and you need first calculate the share of each firm.

### Example:

```{r, eval=FALSE}
df_total <- df[, c(1,2)] # select all rows and the column that contains the firm or firm id and the values (e.g. total sales of a market or submarket)

share_df_total <- df_total %>%
  mutate(share = total/sum(total)) %>% # creates a new column with calculated market share of each firm
  group_by(firm_id)
share_df_total # should give you a tibble with 3 columns, 
```

Then just go ahead with:

```{r, eval=FALSE}
con_share_df_total <- concstats(share_df_total) # calculates all measures
con_share_df_total # resulting in a table with 12 measures
```

### Note:

It should be clear that any vector of market shares should sum up to 1 - in other words, in decimal form. 
