---
title: "Getting started"
description: >
  Learn what concstats is about
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## <i class="far fa-lightbulb"></i> What is concstats?

concstats is a small package that helps you to **calculate the structure,
concentration or diversity measures of a given market.**  
You can reach that goal in different ways:  
a) calculating an individual measure, e.g. the Herfindahl Hirschman Index (HHI)
   or its dual.  
b) calculating a group of measures, e.g. structural measures, or  
c) a selected set of measures in a one-step procedure.


## <i class="fas fa-download"></i> How can you install it?

You can install `concstats` directly from CRAN

```{r, eval=FALSE}
install.packages("concstats") # Market Structure, Concentration, and Inequality
                              # Measures
library(concstats)

```

To install the latest development version of `concstats` directly from [GitHub](https://github.com/schneiderpy/concstats) or use: 

```{r, message=FALSE, warning=FALSE, results = "hide", eval=FALSE}
library(devtools)
devtools::install_github("schneiderpy/concstats")
```

# </a><i class="fas fa-table"></i> Why did I build it?

Measures of concentration and competition are important and give a first insight
of a given market structure in a particular market. 
**They are important to determine public policies and strategic corporate  
decisions**. However, in research and in practice the most commonly used measure  
is the Herfindahl Hirschman Index. The `concstats` package offers a set of 
alternative measures.

`concstats` allows you to **calculate quickly a particular measure or a group of
measures to give you a better overview of a given market situation and therefore
reducing uncertainty.** **It can be used by practitioners and academics alike.**

## <i class="fas fa-code"></i> How can it be used?
```{r, echo=FALSE, eval=FALSE}
#' @srrstats {EA1.0} identifying target audience on line#60
#' @srrstats {EA1.2} scope of the package on line#61-62
```
`concstats` should be used by practitioners and academics **who are concerned
about structure, concentration, diversity or inequality** in general or on a
regular basis. **A single standard measure like the Herfindahl Hirschman Index
(HHI) may lead to erroneous conclusions.
** The `concstats` package offers a straightforward way to calculate standard
and alternative measures** given a data set of market participants and their
participation (in relative values) in a particular market at a given point of
time or over time.
concstats provides currently the following basic functions:

- **concstats_concstats** calculates eight pre-selected measures in a one-step
procedure to provide a first insight of the market.
- **concstats_mstruct** is a wrapper for the following measures: **firm**,
**nrs_eq** (Numbers equivalent), **top**, **top 3** and **top 5** market share,
and **all** calculates all measures as a group. Each of the measures within a
group can be accessed directly using the prefix **concstats_**, e.g.
**concstats_firm**.
- **concstats_comp** is a group wrapper for the following concentration measures:
**hhi**, the **dual of the hhi**, the **min. of the hhi**, the
**dominance index**, the **stenbacka index**, and finally **all** which
calculates all measures as a group.
- **concstats_inequ** is a wrapper for inequality and diversity measures and
contains: **entropy**, **gini coefficient**, **simpson index**, the
**palma ratio** and the alternative **grs** measure, **all** calculates the
group measures.
